---
title: "Rain classification example"
author: "C Devenish & O Metcalfe"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{example_rain_classification}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(hardRain)
```


# 1. Obtain wav filenames

NOTE: this will download 100 15s wav files (120 MB) to a new directory created in your home directory.

Create temporary directories for the rain and test files
```{r}
dir.create(tmp1 <- tempfile("rainBR_", tmpdir = getwd()))
dir.create(tmp2 <- tempfile("testBR_", tmpdir = getwd()))
```

Download the rain and test data:
* rain data: 100 wav files known to be hard rain
* test data: 100 wav files (50 rain, 50 non-rain) 

```{r}
download.file(url = xxx, destfile = tmp1)
 
# get the filenames (using regular expressions to differentiate the filenames)
train.BR  <- list.files(path = tmp1, pattern = "BR.*rain.*\\.wav$")
train.MCR <- list.files(path = tmp1, pattern = "MCR.*rain.*\\.wav$")
test.BR <- list.files(path = tmp1, pattern = "BR.*test.*\\.wav$")
test.MCR <- list.files(path = tmp1, pattern = "MCR.*test.*\\.wav$")
```

Calculate the threshold using default settings - for two frequency bands
```{r}

# devtools::install_github("https://github.com/Cdevenish/hardRain")
# library(hardRain)

trBR <- getThreshold(train.BR, fn = "spec")
trBR

trMCR <- getThreshold(train.MCR, fn = "spec")
trMCR
```


Inspect the PSD and signal to noise values for rain and non-rain files at specific frequency bands

```{r}
metBR <- getMetrics(test.BR, fn = "spec")
metMCR <- getMetrics(test.MCR, fn = "spec")

# bind into a data frame and add location
psd_s2n <- data.frame(filename = c(rownames(metBR), rownames(metMCR)), 
                      rbind(metBR, metMCR), 
                      loc = c(rep("BR", length(test.BR)), rep("MCR", length(test.MCR))))

# add test status
psd_s2n$rain <- ifelse(grepl("rain", psd_s2n$filename), TRUE, FALSE)

# Draw a boxplot to see separation between rain and non-rain for each metric

boxplot(psd ~ rain + loc, data = psd_s2n)
boxplot(s2n ~ rain + loc, data = psd_s2n)

## Or with ggplot


```


Classify the test files using the thresholds obtained above

```{r}
resBR <- classifyRain(test.BR, t.values = trBR, fn = "spec", threshold = "min")
resMCR <- classifyRain(test.MCR, t.values = trMCR, fn = "spec", threshold = "min")

# How many files identified as rain/non-rain?
# table(resBR)
```

